---
title: "transform.hazards"
author: "Pål Ryalen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transform.hazards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Transforms cumulative hazard estimates to estimate survival analysis parameters specified by differential equations. We demonstrate how the method works on some commonly studied parameters.


## Parameters

We are interested in assessing parameters that solve differential equations driven by cumulative hazards $A$, i.e.
\begin{equation}X_t = X_0 + \int_0^t F(X_s) dA_s,\end{equation}
where $F = (F_1,F_2,\cdots)$ is Lipschitz, two times continuously differentiable, and satisfies a linear growth bound. Several examples of such parameters can be found in [^fn1] [^fn2]. The main function *pluginEstimate* in this package estimate such parameters. Among other things, *pluginEstimate* has the following input

- The increments of the cumulative hazard estimates
- The function $F$
- The Jacobian matrices of the columns of $F$, i.e. $J_{F_1},J_{F_2},\cdots$, as a list

We illustrate how to provide the correct inputs by use of examples below.


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)

library(timereg)
library(transform.hazards)
```

## Survival

The survival function $S$ solves a differential equation with initial value 1;
\begin{equation} S_t = 1 - \int_0^t S_{s} dA_s.\end{equation}

We generate a sample of exponentially distributed survival times with independent censoring, and calculate cumulative hazard estimates

```{r, fig.show='hold',results = "hide",echo=T}
n1 <- 300
Samp1 <- rexp(n1,1)
cTimes <- runif(n1)*4
Samp1 <- pmin(Samp1,cTimes)
isNotCensored <- cTimes != Samp1
startTimes1 <- rep(0,n1)

aaMod1 <- aalen(Surv(startTimes1,Samp1,isNotCensored)~1)
```
We want to estimate the survival function. The cumulative hazard estimates are inserted as a time-ordered matrix where each column is an increment of the vector of cumulatie hazard estimates;

```{r, fig.show='hold',results = "hide",echo=T}
dA_est1 <- diff(c(0,aaMod1$cum[,2]))
hazMatrix <- matrix(dA_est1,nrow=1)
```

In this case, $F$ takes the simple form $F(x) = -x$, and its Jacobian is therefore $J_{F}(x) = -1$. These must be provided as matrix-valued functions;

```{r, fig.show='hold',results = "hide",echo=T}
F_survival <- function(X)matrix(-X,nrow=1,ncol=1)
F_survival_JacobianList <- list(function(X)matrix(-1,nrow=1,ncol=1))
```

We obtain plugin estimates of $S$ and its covariance using the call

```{r, fig.show='hold',results = "hide",echo=T}
param <- pluginEstimate(n1, hazMatrix, F_survival, F_survival_JacobianList,matrix(1,nrow=1,ncol=1),matrix(0,nrow=1,ncol=1))
```
 Finally, we may plot the results with approximate 95 % confidence intervals
 
 
```{r, fig.show='hold', echo=T}
times1 <- aaMod1$cum[,1]
plot(times1,param$X,type="s",xlim=c(0,4),xlab="t",ylab="",main="Survival")
lines(times1,param$X + 1.96 * sqrt(param$covariance[1,1,]),type="s",lty=2)
lines(times1,param$X - 1.96 * sqrt(param$covariance[1,1,]),type="s",lty=2)
lines(times1,exp(-times1),col=2)
legend("topright",c("estimated","true"),lty=1,col=c(1,2),bty="n")
```

## Restricted mean survival

The restricted mean survival function $R$ solves the system
\begin{equation} \begin{pmatrix} S_t \\ R_t \end{pmatrix} = \begin{pmatrix}1 \\ 0 \end{pmatrix} + \int_0^t \begin{pmatrix} -S_{s} & 0 \\ 0 & S_s \end{pmatrix}d \begin{pmatrix}A_s \\ s \end{pmatrix},\end{equation}

where $S$ is the survival function. Here, the colums of $F$, $F_1$ and $F_2$, are given by
\begin{equation} F_1(x_1,x_2) = \begin{pmatrix} -x_1 \\ 0 \end{pmatrix}, F_2(x_1,x_2) = \begin{pmatrix} 0 \\ x_1 \end{pmatrix}.\end{equation}
The Jacobian matrices are therefore
\begin{equation} J_{F_1}(x_1,x_2) = \begin{pmatrix} -1 & 0 \\ 0 & 0 \end{pmatrix}, J_{F_2}(x_1,x_2) = \begin{pmatrix} 0 & 0 \\ 1 & 0 \end{pmatrix}. \end{equation}
We define these along with initial values below
```{r, fig.show='hold',results = "hide",echo=T}
F_restrict <- function(X)matrix(c(-X[1],0,0,X[1]),nrow=2)
F_restrict_JacobianList <- list(function(X)matrix(c(-1,0,0,0),nrow=2),
                                function(X)matrix(c(0,0,1,0),nrow=2,byrow=T))

X0_restrict <- matrix(c(1,0),nrow=2)
V0_restrict <- matrix(0,nrow=2,ncol=2)
```

The restricted mean survival is a 'regular' (i.e. Lebesgue) integral, and we must therefore provide the time increments. We choose the time interval $[0,4]$ in $10^4$ increments:

```{r, fig.show='hold',results = "hide",echo=T}
fineTimes <- seq(0,4,length.out = 1e4+1)

tms <- sort(unique(c(fineTimes,times1)))

hazMatrix <- matrix(0,nrow=2,ncol=length(tms))
hazMatrix[1,match(times1,tms)] <- dA_est1
hazMatrix[2,] <- diff(c(0,tms))
```

We obtain plugin estimates using the call (note the last argument that in this example can be used to improve efficiency);

```{r, fig.show='hold',results = "hide",echo=T}
param <- pluginEstimate(n1, hazMatrix, F_restrict, F_restrict_JacobianList,X0_restrict,V0_restrict,isLebesgue = 2)
```

We plot the results
```{r, fig.show='hold', echo=T}
plot(tms,param$X[2,],type="s",xlim=c(0,4),ylim=c(0,1.5),xlab="t",ylab="",main="Restricted mean survival")
lines(tms,param$X[2,] + 1.96 * sqrt(param$covariance[2,2,]),type="s",lty=2)
lines(tms,param$X[2,] - 1.96 * sqrt(param$covariance[2,2,]),type="s",lty=2)
lines(tms,1 - exp(-tms),col=2)
legend("topleft",c("estimated","true"),lty=1,col=c(1,2),bty="n")

```


## Relative survival

We generate another set of exponentially distributed survival times and compare the two groups. A new matrix of cumulative hazard increments must be created;

```{r, fig.show='hold',results = "hide",echo=T}
n2 <- 200
Samp2 <- rexp(n2,1.3)
censTimes2 <- runif(n2)*3
Samp2 <- pmin(Samp2,censTimes2)
isNotCensored2 <- censTimes2 != Samp2
startTimes2 <- rep(0,n2)

aaMod2 <- aalen(Surv(startTimes2,Samp2,isNotCensored2)~1)
dA_est2 <- diff(c(0,aaMod2$cum[,2]))

times2 <- aaMod2$cum[,1]
times <- sort(unique(c(times1,times2)))

hazMatrix <- matrix(0,nrow=2,ncol=length(times))
hazMatrix[1,match(times1,times)] <- dA_est1
hazMatrix[2,match(times2,times)] <- dA_est2
```
The relative survival function $RS$ between groups 1 and 2 solve the equation
\begin{equation} RS_t = 1 + \int_0^t \begin{pmatrix} -RS_s & RS_s\end{pmatrix} d\begin{pmatrix} A_s^1 \\ A_s^2 \end{pmatrix}, \end{equation}

i.e. $F(x) = (-x,x)$. The Jacobians of the columns of $F$ are $J_{F_1}(x) = -1$, and $J_{F_2}(x) = 1$. We specify these functions as follows:

```{r, fig.show='hold',results = "hide",echo=T}
F_relsurv <- function(X)matrix(c(-X,X),nrow=1)
F_relsurv_JacobianList <- list(function(X)matrix(-1,nrow=1),
                               function(X)matrix(1,nrow=1))
```

The estimates are then obtained by the call

```{r, fig.show='hold',results = "hide",echo=T}
param <- pluginEstimate(n1+n2, hazMatrix, F_relsurv, F_relsurv_JacobianList,matrix(1,nrow=1,ncol=1),matrix(0,nrow=1,ncol=1))
```
 
 We may plot the results with approximate 95% confidence intervals
 
 
```{r, fig.show='hold', echo=T}
plot(times,param$X,type="s",xlim=c(0,4),ylim=c(-0.5,4),xlab="t",ylab="",main="Relative survival")
lines(times,param$X + 1.96 * sqrt(param$covariance[1,1,]),type="s",lty=2)
lines(times,param$X - 1.96 * sqrt(param$covariance[1,1,]),type="s",lty=2)
lines(times,exp(-times*(1/1.3 - 1)),col=2)
legend("topleft",c("estimated","true"),lty=1,col=c(1,2),bty="n")

```

## Cumulative incidence
We may be in a situation with two competing risks. The cumulative incidences $C^1,C^2$ solve the system
\begin{equation} \begin{pmatrix} S_t \\ C_t^1 \\ C_t^2 \end{pmatrix} = \begin{pmatrix} 1 \\ 0 \\ 0 \end{pmatrix} + \int_0^t \begin{pmatrix} -S_s & -S_s \\ S_s & 0 \\ 0 & S_s \end{pmatrix} d \begin{pmatrix} A^1_s \\ A^2_s \end{pmatrix}, \end{equation}
where $S$ is the survival. The first columns of $F$ are

\begin{equation}F_1(x_1,x_2,x_2) = \begin{pmatrix} -x_1 \\ x_1 \\ 0 \end{pmatrix},  F_2(x_1,x_2,x_2) = \begin{pmatrix} -x_1 \\ 0 \\ x_1 \end{pmatrix} \end{equation}

The reader may verify that the Jacobian matrix of $F_1$ and $F_1$ are
\begin{equation} J_{F_1}(x_1,x_2,x_3) = \begin{pmatrix} -1 & 0 & 0 \\ 1 & 0 & 0 \\ 0 & 0 & 0 \end{pmatrix} , J_{F_2}(x_1,x_2,x_3) = \begin{pmatrix} -1 & 0 & 0 \\ 0 & 0 & 0 \\ 1 & 0 & 0 \end{pmatrix} \end{equation}

We specify the function $F$ and the list of the Jacobian matrices below, along with the initial values:
```{r, fig.show='hold',results = "hide",echo=T}
F_cuminc <- function(X)matrix(c(-X[1],-X[1],X[1],0,0,X[1]),nrow=3,byrow=T)
F_cuminc_JacobianList <- list(function(X)matrix(c(-1,0,0,1,0,0,0,0,0),nrow=3,byrow=T),
                               function(X)matrix(c(-1,0,0,0,0,0,1,0,0),nrow=3,byrow=T))
X0_cuminc <- matrix(c(1,0,0),nrow=3)
v0_cuminc <- matrix(0,nrow=3,ncol=3)
```

Now for obtaining hazard estimates. We generate survival times first, before sampling two causes of death (and censoring). We then estimate the cause-specific cumulative hazards, and transform the estimates:

```{r, fig.show='hold',results = "hide",echo=T}
# Generate the data
dfr <- data.frame(from=0,to=rexp(n1+n2,1),from.state = 1)

# Adding causes of death (to.state = 2 or 3) and censoring (to.state = 0)
dfr$to.state <- sample(c(0,2,3),n1+n2,replace=T,prob=c(0.2,0.2,0.6))


# Obtaining cause-specific cumulative hazard estimates and extracting the increments
aamod22 <- aalen(Surv(from,to,to.state %in% c(2)) ~1, data=dfr)
aamod33 <- aalen(Surv(from,to,to.state %in% c(3)) ~1, data=dfr)
tmss = sort(unique(c(aamod22$cum[,1],aamod33$cum[,1])))
hazMatrix = matrix(0,nrow=2,ncol=length(tmss))
mt1 = match(aamod22$cum[,1],tmss)
mt2 = match(aamod33$cum[,1],tmss)
hazMatrix[1,mt1] = diff(c(0, aamod22$cum[,2] ))
hazMatrix[2,mt2] = diff(c(0, aamod33$cum[,2] ))

# Transforming the estimates
param <- pluginEstimate(n1+n2,hazMatrix,F_cuminc,F_cuminc_JacobianList,X0_cuminc,v0_cuminc)
```

Finally, we plot the estimates of the cumulative incidences $C^1$ and $C^2$ with confidence intervals:

```{r, fig.show='hold', echo=T}
plot(tmss,param$X[2,],type="s",xlim=c(0,4),ylim=c(0,1),xlab="t",ylab="",main="Cumulative incidence")
lines(tmss,param$X[2,] + 1.96 * sqrt(param$covariance[2,2,]),type="s",lty=2)
lines(tmss,param$X[2,] - 1.96 * sqrt(param$covariance[2,2,]),type="s",lty=2)
lines(tmss,param$X[3,],type="s",col=3)
lines(tmss,param$X[3,] + 1.96 * sqrt(param$covariance[3,3,]),type="s",lty=2,col=3)
lines(tmss,param$X[3,] - 1.96 * sqrt(param$covariance[3,3,]),type="s",lty=2,col=3)
legend("topleft",c(expression(C^1),expression(C^2)),lty=1,col=c(1,3),bty="n")
```



## Restricted mean survival -- comparing two groups

By re-specifying the ODE system we can compare two groups. We illustrate this in the restricted mean survival example. Consider the system

\begin{equation} \begin{pmatrix} R_t^1 \\ R^2 \\ RD \\ S^1 \\ S^2 \end{pmatrix} = \begin{pmatrix} 0 \\ 0 \\ 0 \\ 1 \\ 1 \end{pmatrix} + \int_0^t \begin{pmatrix} S^1_{s} & 0 & 0 \\ S^2_{s} & 0 & 0 \\ S^1_{s} - S^2_{s} & 0 & 0 \\ 0 & -S_s^1 & 0 \\ 0 & 0 & -S_s^2 \end{pmatrix}d \begin{pmatrix}s \\ A_s^1 \\ A_s^2 \end{pmatrix},\end{equation}

where $S^i$ is the survival function in group $i$. 
<!-- Here, the colums of $F$, $F_1$ and $F_2$, are given by -->
<!-- \begin{equation} F_1(x_1,x_2) = \begin{pmatrix} -x_1 \\ 0 \end{pmatrix}, F_2(x_1,x_2) = \begin{pmatrix} 0 \\ x_1 \end{pmatrix}.\end{equation} -->
<!-- The Jacobian matrices are therefore -->
<!-- \begin{equation} J_{F_1}(x_1,x_2) = \begin{pmatrix} -1 & 0 \\ 0 & 0 \end{pmatrix}, J_{F_2}(x_1,x_2) = \begin{pmatrix} 0 & 0 \\ 1 & 0 \end{pmatrix}. \end{equation} -->
We specify matrix-valued functions and initial values as before:

```{r, fig.show='hold', echo=T}
F_restrict_diff <- function(X)matrix(c(X[4],0,0,
                                       X[5],0,0,
                                       X[4]-X[5],0,0,
                                       0,-X[4],0,
                                       0,0,-X[5]),nrow=5,byrow=T)
F_restrict_diff_JacobianList <- list(function(X)matrix(c(0,0,0,1,0,
                                                         0,0,0,0,1,
                                                         0,0,0,1,-1,
                                                         rep(0,10)),nrow=5,byrow=T),
                                     function(X)matrix(c(rep(0,18),-1,rep(0,6)),nrow=5,byrow=T),
                                     function(X)matrix(c(rep(0,24),-1),nrow=5,byrow=T))

X0_restrict_diff <- matrix(c(0,0,0,1,1),nrow=5)
V0_restrict_diff <- matrix(0,nrow=5,ncol=5)
```


We compare RMS for males and females in the *flchain* data set in the *survival* package. Specifying *hazmatrix*:

```{r, fig.show='hold', echo=T}

aa_male = aalen(Surv(futime,death==1)~1, data = flchain[flchain$sex=="M",]) 
aa_female = aalen(Surv(futime,death==1)~1, data = flchain[flchain$sex=="F",]) 

tms <- sort(unique(c(fineTimes,aa_male$cum[,1], aa_female$cum[,1])))
mt_male = match(aa_male$cum[,1],tms)
mt_female = match(aa_female$cum[,1],tms)

hazMatrix <- matrix(0,nrow=3,ncol=length(tms))
hazMatrix[1,] <- diff(c(0,tms))
hazMatrix[2,mt_male] <- diff(c(0,aa_male$cum[,2]))
hazMatrix[3,mt_female] <- diff(c(0,aa_female$cum[,2]))
  
  
plugEst <- pluginEstimate(1, hazMatrix, F_restrict_diff, F_restrict_diff_JacobianList,X0_restrict_diff,V0_restrict_diff,isLebesgue = 1)
param = list(plugEst=plugEst, tms=tms)
```

Plotting $\hat{RD} = \hat R^1 - \hat R^2$ with 95\% confidence intervals:

```{r, fig.show='hold', echo=T}
ylims = c(1.1*min(param$plugEst$X[3,] - 1.96 * sqrt(param$plugEst$covariance[3,3,])),
          1.1*max(param$plugEst$X[3,] + 1.96 * sqrt(param$plugEst$covariance[3,3,])))
plot(param$tms,param$plugEst$X[3,],type="s",ylim=ylims,ylab="",main="RMS difference",xlab="years")
lines(param$tms,param$plugEst$X[3,] + 1.96 * sqrt(param$plugEst$covariance[3,3,]),type="s",lty=2)
lines(param$tms,param$plugEst$X[3,] - 1.96 * sqrt(param$plugEst$covariance[3,3,]),type="s",lty=2)
abline(a=0,b=0,col=2)
```




## Cumulative incidence -- comparing two groups


Consider a similar example with differences of cumulative incidence functions

\begin{equation} \begin{pmatrix} S^a_t \\ S_t^b \\ C_t^{a,1} \\ C_t^{a,2} \\ C_t^{b,1} \\ C_t^{b,2} \\ CD_t^1 \\ CD_t^2 \end{pmatrix} = \begin{pmatrix} 1 \\ 1 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{pmatrix} + \int_0^t \begin{pmatrix} -S_s^a & -S_s^a & 0 & 0 \\ 0 & 0 & -S_s^b & -S_s^b \\ S_s^a & 0 & 0 & 0 \\ 0 & S_s^a & 0 & 0 \\ 0 & 0 & S_s^b  & 0 \\ 0 & 0 & 0 & S_s^b \\ S_s^a & 0 & -S_s^b & 0 \\ 0 & S_s^a & 0 & -S_s^b  \end{pmatrix} d \begin{pmatrix} A^{a,1}_s \\ A^{a,2}_s \\ A^{b,1}_s \\ A^{b,2}_s \end{pmatrix}, \end{equation}
where $S^a$ is the survival in group $a$ and $S^b$ is the survival, and $C^{a,j}$ is the cumulative incidence for cause $j$ in group $a$ and similarly for group $b$. $CD^i = C^{a,i} - C^{b,i}$ is the cumulative incidence difference for cause $i$.


```{r, fig.show='hold', echo=T}
F_cuminc_diff <- function(X)matrix( c(-X[1],-X[1],0,0,
                                      0,0,-X[2],-X[2],
                                      X[1],0,0,0,
                                      0,X[1],0,0,
                                      0,0,X[2],0,
                                      0,0,0,X[2],
                                      X[1],0,-X[2],0,
                                      0,X[1],0,-X[2]) ,nrow=8,byrow=T)
F_cuminc_diff_JacobianList <- list(function(X)matrix(c(-1, rep(0,7),
                                                       rep(0,8),
                                                       1, rep(0,7),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       1, rep(0,7),
                                                       rep(0,8)),nrow=8,byrow=T),
                                   function(X)matrix(c(-1, rep(0,7),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       1, rep(0,7),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       1, rep(0,7)),nrow=8,byrow=T),
                                   function(X)matrix(c(rep(0,8),
                                                       0,-1, rep(0,6),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       0,1, rep(0,6),
                                                       rep(0,8),
                                                       0,-1, rep(0,6),
                                                       rep(0,8)),nrow=8,byrow=T),
                                   function(X)matrix(c(rep(0,8),
                                                       0,-1, rep(0,6),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       rep(0,8),
                                                       0,1, rep(0,6),
                                                       rep(0,8),
                                                       0,-1, rep(0,6)),nrow=8,byrow=T))


X0_cuminc_diff <- matrix(c(1,1,0,0,0,0,0,0),nrow=8)
v0_cuminc_diff <- matrix(0,nrow=8,ncol=8)
```





We compare cumulative incidences for recurrence vs death between two treatment groups "Lev" and "Lev+5FU" in the *colon* data set. Specifying *hazmatrix*:

```{r, fig.show='hold', echo=T}

fr_lev <- colon[colon$rx == "Lev",]
fr_lev$time <- fr_lev$time /365
fr_lev5u <- colon[colon$rx == "Lev+5FU",]
fr_lev5u$time <- fr_lev5u$time /365

fit_rec_lev = aalen(Surv(time,etype==1 & status==1)~1,data=colon[colon$rx == "Lev",])
fit_death_lev = aalen(Surv(time,etype==2 & status==1)~1,data=colon[colon$rx == "Lev",])

fit_rec_lev5u = aalen(Surv(time,etype==1 & status==1)~1,data=colon[colon$rx == "Lev+5FU",])
fit_death_lev5u = aalen(Surv(time,etype==2 & status==1)~1,data=colon[colon$rx == "Lev+5FU",])

tms2 = sort(unique(c(fit_rec_lev$cum[,1],fit_rec_lev5u$cum[,1],
                     fit_death_lev$cum[,1],fit_death_lev5u$cum[,1])))
        
dA_rec_lev = dA_rec_lev5u = dA_death_lev = dA_death_lev5u = rep(0, length(tms2))
  
  dA_rec_lev[match(fit_rec_lev$cum[,1], tms2)] = diff(c(0,fit_rec_lev$cum[,2]))
  dA_death_lev[match(fit_death_lev$cum[,1], tms2)] = diff(c(0,fit_death_lev$cum[,2]))
  
  dA_rec_lev5u[match(fit_rec_lev5u$cum[,1], tms2)] = diff(c(0,fit_rec_lev5u$cum[,2]))
  dA_death_lev5u[match(fit_death_lev5u$cum[,1], tms2)] = diff(c(0,fit_death_lev5u$cum[,2]))
  
  hazMatrix <- matrix(0,nrow=4,ncol=length(tms2))
  hazMatrix[1,] <- dA_rec_lev
  hazMatrix[2,] <- dA_death_lev
  hazMatrix[3,] <- dA_rec_lev5u
  hazMatrix[4,] <- dA_death_lev5u

plugEst <- pluginEstimate(1, hazMatrix, F_cuminc_diff, F_cuminc_diff_JacobianList,X0_cuminc_diff,v0_cuminc_diff)
param = list(plugEst=plugEst, tms=tms2)
```

Plotting $\hat{CD}^1$ and $\hat{CD}^2$ with 95\% confidence intervals:

```{r, fig.show='hold', echo=T}
ymxx = max(param$plugEst$X[7,] + 1.96 * sqrt(param$plugEst$covariance[7,7,]),
           param$plugEst$X[8,] + 1.96 * sqrt(param$plugEst$covariance[8,8,]))
ymnn = min(param$plugEst$X[7,] - 1.96 * sqrt(param$plugEst$covariance[7,7,]),
           param$plugEst$X[8,] - 1.96 * sqrt(param$plugEst$covariance[8,8,]))

plot(param$tms/365, param$plugEst$X[7,],type="s",ylim=c(ymnn,ymxx),main="Cumulative incidence difference for recurrence",
    ylab="",xlab="Years")
lines(param$tms/365, param$plugEst$X[7,] + 1.96 * sqrt(param$plugEst$covariance[7,7,]),type="s",lty=2)
lines(param$tms/365, param$plugEst$X[7,] - 1.96 * sqrt(param$plugEst$covariance[7,7,]),type="s",lty=2)
abline(a=0,b=0,col=2)



plot(param$tms/365, param$plugEst$X[8,],type="s",ylim=c(ymnn,ymxx),main="Cumulative incidence difference for death",
     ylab="",xlab="Years")
lines(param$tms/365, param$plugEst$X[8,] + 1.96 * sqrt(param$plugEst$covariance[8,8,]),type="s",lty=2)
lines(param$tms/365, param$plugEst$X[8,] - 1.96 * sqrt(param$plugEst$covariance[8,8,]),type="s",lty=2)
abline(a=0,b=0,col=2)
```








  




[^fn1]: [Transforming cumulative hazard estimates] (Bioaetrika, 2018).
[^fn2]: [On null hypothesis in survival analysis] (Biometrics, 2019) 
